### 2021-02-16 text/abstract/abstract.tex
tex structure

\hypertarget{abstract}{%
	\section*{Abstract}\label{abstract}}

Lorem ipsum.