### 2021-03-09 text/abstract/abstract.tex
update conclusion

\hypertarget{abstract}{%
	\section*{Abstract}\label{abstract}}

Lorem ipsum.

\newpage

\emph{"Never spend 6 minutes doing something by hand when you can spend 6 hours failing to automate it."}

Zhuowei Zhang