### 2021-03-12 text/titlepage/titlepage.tex
add conclusion

\newpage
%\sffamily
\thispagestyle{empty}
\large
\begin{center}
	\includegraphics[width=5cm]{./logo.png}\\[.5cm]
	Universität Zürich, Philosophische Fakultät\\[.25cm]
 	Zentralbibliothek Zürich, MAS Bibliotheks- und Informationswissenschaft\\[1cm]
	\hrule
	\vspace{2cm}
	\flushleft
	\textbf{\LARGE Machine indexing of institutional repositories: indexing Edoc with Annif as proof of concept}\\[2cm]
	Dr.\ Maximilian G.\ Hindermann \orcidicon{0000-0002-9337-4655}\\
	Spalentorweg 46, CH-4051 Basel\\
	+41 79 277 0612\\
	\href{mailto:maximilian.hindermann@gmail.com}{maximilian.hindermann@gmail.com}\\
	\vspace{1cm}
	\flushleft
	Silke Bellanger, Referentin \orcidicon{0000-0003-1952-1933}\\[.25cm]
	Dr.\ Andreas Ledl, Ko-Referent \orcidicon{0000-0002-0629-0446}\\
	%Prof.\ Marco J.\ Nathan\\[.25cm]
	%\begin{minipage}{.4\textwidth}
	%\end{minipage}
	\vfill
	\center
	1.\ April 2021
\end{center}
\normalfont
\normalsize