### 2021-03-15 text/abstract/abstract.tex
add abstract

\hypertarget{abstract}{%
	\section*{Abstract}\label{abstract}}

The potential of some institutional repositories is hampered by the lack of subject indexing. I argue that this situation can be amended with the help of machine indexing. As proof of concept I show that Edoc --- the University of Basel's repository --- can be successfully indexed using the Annif-client Python library. In order to do so, I assess the performance of hundreds of Annif configurations in assigning subject terms to a sample data set against a gold standard constructed from cleaned and reconciled author keywords.

\newpage

\emph{"Never spend 6 minutes doing something by hand when you can spend 6 hours failing to automate it."}

Zhuowei Zhang